---
title: "week1-discussion"
author: "Caitlin Nordheim-Maestas"
date: "2024-04-02"
output: html_document
---

This week’s Discussion Section
In this course, we will work a lot with writing and applying functions; You’ve all been exposed to writing functions in previous courses - but just to make sure that everyone remembers the basic syntax in R, complete the following:

a. Write a function that computes energy produced from a photovoltaic system if you know the average annual solar radiation.  You can use the following formula: 

    $E = A * r * H * PR$

    * E is energy (kWh),
    * A is the solar panel area (m2)
    * r is panel yield (0-1) (manufacture efficiency - usually around 0.2),
    * PR is performance ratio (0-1) (accounting for site factors that impact efficiency usually around 0.75) 
    * H is annual average solar radiation (kWh)


b. Write your function in R; make sure that you add documentation; include default values for *r* and *PR* in the function definition; save the function in a single file to upload to Canvas.

# Answer

```{r}
# creating a function that outputs "E", the energy, in kWh
# parameters and units: A: area of solar panel m2, r: panel yeild in manufacture efficiency, PR: performance ratio, H: annual average solar radiation (kWh)
energy <- function(A, r=0.2, H, PR=0.75){ 
  # specifying inputs of the function, here they are the parameters of the model
  # note: r and PR are assigned default values by assigning them values in the function() argument, but we can always assign something else manually if we choose by saying r= when we use the function
  E = A * r * H * PR  # E is the solution to multiplying all of the other parameters together
  return(E) # print out the value of the energy in kWh
}

```
# Testing the function:

```{r}
# let's try it!
energy(A=10, H=1) # only specify the parameters without a default

# let's confirm it works by hard coding it
10 * 0.2 * 1 * 0.75 # same answer!
```

