---
title: "Week-2 homework: Almond yeild"
format: html
editor: 
  markdown: 
    wrap: 72
---

# ASSIGNMENT 2 Almond Yields

Lobell et al., 2006 used data to build models of tree crop yield for
California that captured how climate variation (place and time) might
influence yield - Yield anomaly - Climate variables

Assignment For this assignment you will work in pairs Your goal is to
implement a simple model of almond yield anomaly response to climate -
Inputs: daily times series of minimum, maximum daily temperatures and
precipitation - Outputs: maximum, minimum and mean yield anomoly for a
input time series The Lobell et al. 2006 paper will be the source for
the transfer function that you will use in your model; specifically look
at the equations in table 2.

What you will do - Model Set Up 1. Draw diagram to represent your
model - how it will translate inputs to outputs, with parameters that
shape the relationship between inputs an outputs - on your diagram list
what your inputs, parameters and outputs are with units

2.  Implement your diagram as an R function

Here are some ideas to think though. The climate data is going to need
to be processed (e.g note that the model uses climate data from a
particular month - you have ALL of the climate as daily data). Here are
two possible model outlines to follow

-   Almond_model \<- function(clim_var1, clim_var2, parameters){……}
-   Almond_model \<- function(clim, parameters){……}

The first example is where the climate variables are separately input
into the function - climate data is processed beforehand as part of
model set up

The second is where a climate data frame is the input to the function
and you extract the useful data from it as part of the model. You can
pick which option you prefer (or try both) If you want a coding
challenge, you could create a model that works for any crop

## Run your model for the clim.txt data that is posted on Canvas,

clim.txt has the following columns these 4 columns tell you when climate
observations were made day month year wy (water year)

the next 3 columns are the climate observations maximum (tmax_c) daily
temperature minimum (tmin_c) daily temperature precipitation (precip) on
that day in mm

## check your work

I will tell you that 

-   the **maximum** almond yield anomaly should be approximately 1920 ton/acre 
-   the **lowest** almond yield anomaly should be approximately -0.027 ton/acre 
-   the **mean** almond yield anomaly should be approximately 182 ton/acre

## Notes

We will build on this in next Thursday’s class There are always multiple
ways to code something in R; Remember that we want to strive for our
code being as simple and streamline as possible. Style counts. Make sure
you choose meaningful variable names and add comments. Include comments
at the top of the function to tell the user what the inputs/outputs are
and their units and format.

## What to hand in on Canvas

Two separate files (or a link to a repository with these files) 1. a R
file that has your function definition 2. a Rmarkdown file that includes
your conceptual model (embed as picture) and shows how you applied the
function to the test climate data (clim.txt)

# Import and clean data

```{r}
library(tidyverse)
library(here)
```

NB ran this in terminal to make txt to csv: "mv clim.txt clim.csv"

```{r}
# read in data and clean the dataset
clim <- read_csv(here("week-2", "clim.csv")) %>% 
    separate(col = "day month year wy tmax_c tmin_c precip",
           into =c("day", "month", "year", "wy", "tmax_c", "tmin_c", "precip"),
           sep = " ") %>% 
  filter(month ==2) %>% # Feb is only month that matters to almond yeild
  group_by(year) %>% 
  mutate(tmin_c =as.numeric(tmin_c),
         precip =as.numeric(precip))

clim_summ <- read_csv(here("week-2", "clim.csv")) %>% 
    separate(col = "day month year wy tmax_c tmin_c precip",
           into =c("day", "month", "year", "wy", "tmax_c", "tmin_c", "precip"),
           sep = " ") %>% 
  filter(month ==2) %>% # Feb is only month that matters to almond yeild
  group_by(year) %>% 
  mutate(tmin_c =as.numeric(tmin_c),
         precip =as.numeric(precip)) %>% 
  group_by(year) %>% 
  summarize(month_precip = sum(precip),
            min_temp_for_month = min(tmin_c))

```

# Build Function

```{r}
almonds<- function(Tn, P){ 
  # specifying inputs of the function, here they are the parameters of the model
  Y = -0.015*Tn - 0.0046*Tn^2 - 0.07*P + 0.0043*P^2 + 0.28 
  return(max(Y)) # print out the value of the almond yield (in tons/acre)
}

```

# Test function

-   the **maximum** almond yield anomaly should be approximately **1920** ton/acre 
-   the **lowest** almond yield anomaly should be approximately **-0.027** ton/acre 
-   the **mean** almond yield anomaly should be approximately **182** ton/acre


```{r}
almonds(Tn = clim_summ$min_temp_for_month, P = clim_summ$month_precip)

```

